---
title: "Ski Data Preparation"
output: html_notebook
---

Import Sales Data
```{r}
library(readr)
sales_data <- read_csv("~/data/sales_data.csv")
```

Googlesheets will be used for persistent storgae
```{r}
library(googlesheets)
suppressMessages(library(dplyr))
```

Prepare sheet details

```{r}
gap <- gs_key("1-17_HfRFg6guIf3lDs-Lsa6zGmOhKRU5AkNQvUcWnNc", lookup = FALSE, visibility = "private")
```

Adding Season Information
```{r}

sales_data$year <- strtoi(format(sales_data$date, "%Y"))
sales_data$month <- strtoi(format(sales_data$date, "%m"))
sales_data$yrmonth <- strtoi(format(sales_data$date, "%Y%m"))

yrmonth <- unique(sales_data$yrmonth)
season_end <- (unique(sales_data$year) + 1) *100 +5 

sales_data$season <- 0

for (i in 1:nrow(sales_data)){
  for (j in 1:length(season_end)){
    if (sales_data$yrmonth[i] < season_end[j] && sales_data$season[i] == 0){
      sales_data$season[i] <- j
    }
  }
}

```

Add Preliminary sales data
```{r}
 gs_ws_new(gap, 'sales_data', input = sales_data, trim = TRUE)
```

Prepare Data
```{r}
d <- gs_read(gap, ws = "sales_weather_data")
d$day_week <- weekdays(as.Date(d$date))
d<- d[d$season != 0,]
d$DoW <- factor(d$day_week, levels= c("Monday", 
    "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
```

Add Summary Metrics - Profit & Days
```{r}
d$n_rev <- 0
d$p_rev <- 0
d$n_retCost <- 0
d$p_retCost <- 0
d$n_cost <- 0
d$p_cost <- 0

d$n_inst <- 0
d$p_inst <- 0
d$days_open <- 0
d$days_season <- 0
d$days_ostaffed <- 0
d$days_snow <- 0
d$snow_on_grnd_cm[is.na(d$snow_on_grnd_cm)] <- mean(d$snow_on_grnd_cm, na.rm = TRUE)

d$n_rev[d$promotion == 0] <- d$lessons[d$promotion == 0] * 549 
d$p_rev[d$promotion == 1] <- d$lessons[d$promotion == 1] * 499
d$n_retCost[d$promotion == 0] <- (20 - d$lessons[d$promotion == 0]) * 100
d$p_retCost[d$promotion == 1] <- (36 - d$lessons[d$promotion == 1]) * 100
d$n_cost[d$promotion == 0] <- d$lessons[d$promotion == 0] * 549/2
d$p_cost[d$promotion == 1] <- d$lessons[d$promotion == 1] * 499/2

d$n_inst[d$promotion == 0] <- 20
d$p_inst[d$promotion == 1] <- 36
d$days_open[d$lessons > 0] <- 1
d$days_season[d$season > 0] <- 1

d$total_cost <- (d$n_cost + d$p_cost + d$n_retCost + d$p_retCost) 
d$total_revenue <- (d$n_rev + d$p_rev)
d$profit <- d$total_revenue - d$total_cost
d$total_inst <- d$n_inst + d$p_inst
d$days_ostaffed [d$total_inst > d$lessons] <- 1
d$days_snow [d$snow_on_grnd_cm > 15] <- 1
d$mean_temp_c_factor <- 2 * d$mean_temp_c

```

Create summary
```{r}
k <- d %>%
  group_by(season, DoW, month, year, promotion) %>%
  summarise(total_cost = sum(total_cost), 
            n_cost = sum(n_cost), 
            p_cost = sum(p_cost),
            n_retCost = sum(n_retCost),
            p_retCost = sum(p_retCost),
            total_revenue = sum(total_revenue), 
            n_rev = sum(n_rev), 
            p_rev = sum(p_rev),
            profit = sum(profit), 
            total_inst = sum(total_inst), 
            n_inst = sum(n_inst), 
            p_inst = sum(p_inst), 
            lessons = sum(lessons),
            days_open = sum(days_open),
            days_season = sum(days_season, na.rm = TRUE),
            days_snow = sum(days_snow, na.rm=TRUE),
            days_ostaffed = sum(days_ostaffed),
            total_snow_cm = sum(total_snow_cm, na.rm=TRUE),
            snow_on_grnd_cm = sum (snow_on_grnd_cm, na.rm=TRUE),
            mean_temp_c_factor = sum(mean_temp_c_factor, na.rm=TRUE)
  )

```

Upload new sheet
```{r}
gs_ws_new(gap, 'summary', input = k, trim = TRUE)
```
Prepare data for modelling
```{r}
model_data<- d %>%
  group_by(season, DoW, month, year, promotion) %>%
  summarise(lessons = sum(lessons),
            mean_temp_c_factor = sum(mean_temp_c_factor),
            snow_on_grnd_cm = sum(snow_on_grnd_cm),
            total_snow_cm = sum(total_snow_cm),
            days_season = sum(days_season)
)
```

```{r}
library(stats)

m_temp <- mean(model_data$mean_temp_c_factor/ model_data$days_season, na.rm = TRUE)
s_temp <- sd(model_data$mean_temp_c_factor/ model_data$days_season, na.rm = TRUE)
n_temp <- length(model_data$mean_temp_c_factor)
m_snow <- mean(model_data$snow_on_grnd_cm/ model_data$days_season, na.rm = TRUE)
s_snow <- sd(model_data$snow_on_grnd_cm/ model_data$days_season, na.rm = TRUE)
n_snow <- length(model_data$snow_on_grnd_cm)

```

Assess the distribution of Snow on the ground and fitness of normal function
```{r}
h <- hist(model_data$snow_on_grnd_cm/model_data$days_season, breaks = 13)
xhist<-c(min(h$breaks),h$breaks)
yhist<-c(0,h$density,0)
x.norm<-rnorm(n=n_snow,m=m_snow,sd=s_snow)
xfit<-seq(min(x.norm),max(x.norm),length=50)
yfit<-dnorm(xfit,mean=mean(x.norm),sd=sd(x.norm))
plot(xhist,yhist,type='s',ylim=c(0,max(yhist,yfit)), main='Avg. Snow on Ground Normal Fit'
          , xlab = 'Snow in cm', ylab = 'Probability')
lines(xfit,yfit, col='red')

```

```{r}
shapiro.test(model_data$snow_grnd/model_data$days_season)
```

Assess the distribution of Snow on the ground and fitness of normal function
```{r}
h <- hist(model_data$mean_temp_c_factor/model_data$days_season, breaks = 13)
xhist<-c(min(h$breaks),h$breaks)
yhist<-c(0,h$density,0)
x.norm<-rnorm(n=n_temp,m=m_temp,sd=s_temp)
xfit<-seq(min(x.norm),max(x.norm),length=50)
yfit<-dnorm(xfit,mean=mean(x.norm),sd=sd(x.norm))
plot(xhist,yhist,type='s',ylim=c(0,max(yhist,yfit)), main='Avg. Temperature Normal Fit'
     , xlab = 'Temperature in °C', ylab = 'Probability')
lines(xfit,yfit, col='red')

```

```{r}
hist(model_data$lessons, main="Distribution of Lessons", xlab="# Lessons")
```


```{r}
model <- lm(lessons ~ snow_on_grnd_cm + mean_temp_c_factor + DoW + month + promotion + days_season, data = model_data)

```

```{r}
summary(model)
```
Prepare parameters sheet
```{r}
library(broom)
g<-cbind(t(data.frame(model$coefficients)),
         data.frame(m_snow, m_temp, s_snow, s_temp))

gs_ws_new(gap, 'model_parameters', input = g, trim = TRUE)
```


Clear memory
```{r}
rm(list=ls(all=TRUE))
```


```{r}
rsconnect::deployApp('~/Documents/School/Masters/MMAi Queens/Semester 1/MMAI863/Assignments/A1/R App/ski-lodge/')
```

Model Backup
```{r}
 model <- reactive({
    start_Date <- input$season_length[1]
    end_Date <- input$season_length[2]
    
    # Filter days of interest
    x <- model_dates[model_dates$Date >= start_Date,]
    x <- x[x$Date <= end_Date,]
    x$count <- 1
    
    # Summarize all concepts
    y<- x %>%
      group_by(DoW, Month, Year, Day_Num) %>%
      summarise(count = sum(count))
    size <- nrow(y)
    
    # Calculate n = size values from normal dist
    y$snow_avg <- c(rnorm(size, p$m_snow, p$s_snow)) 
    y$temp_avg <- c(rnorm(size, p$m_temp, p$s_temp))
    
    # Calculate total factors
    y$snow <- y$snow_avg * y$count       
    y$temp <- y$temp_avg * y$count 
    
    # Assign factors to week vector
    week <- c(0, p$DoWTuesday, p$DoWWednesday, p$DoWThursday, 
              p$DoWFriday, p$DoWSaturday, p$DoWSunday)
    
    # Calculate lessons
    y$f_int <- as.numeric(p['(Intercept)'])
    y$f_days <- p$days_season * y$Day_Num
    y$f_snow <- p$snow_on_grnd_cm * y$snow
    y$f_temp <- p$mean_temp_c_factor * y$temp
    y$f_month <- p$month * y$Month
    y$f_DoW <- week[y$Day_Num]
    y$lessons <- y$f_int + y$f_days + y$f_snow + y$f_temp +y$f_month +y$f_DoW
    inst <- ceiling(max(y$lessons/y$Day_Num))
    y$profit <- 549/2 * y$lessons - 100*(inst * y$Day_Num - y$lessons)
    y$inst <- inst

    # 'This Season', 'by Month', 'by Week', 'by Weekday'
    group <- grouping()
    
    y <- y %>%
      group_by(!!group) %>%
      summarize_all(mean)
    y
  })
```

```{r}
sample(1:10, 1)

```



